---
title: "Fake ADA PK study"
author: "Monica Gerber"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

I'd like to be able to reproduce the results I get from the Monolix GUI with the lixoftConnectors R package. I use the GUI for exploratory data analysis, but I need to be able to make the results reproducible with a script. 

In this document, I use fake data similar to data we see in some PK studies with anti-drug antibodies (ADA). I run the model with the Monolix API and record all of the settings/options for running that model. Next, I try to reproduce these results with the lixoftConnectors R package by selecting all of the same settings with the R functions it provides.

I compare the results from the GUI and lixoftConnectors package using the diffr R package.

# Data

Data are generated by simulate-data.R.

```{r read-data}
library(tidyverse)
library(here)
library(diffr)
library(lixoftConnectors)

monolix_fake_data <- read_csv(here("data", "monolix-fake-data.csv"))
```

# Monolix GUI

## Setup

* Monolix - 2020R1
* New Project -> Data File -> select "monolix-fake-data.csv"
* ada_model.txt structural model
* Settings -> Project settings
    + Seed: 123456
    + Number of chains
        + Automatic choice: TRUE
        + Computed number of chains: 9 
        + Minimum number of individuals: 50
    + Transition kernels
        + Target acceptance ratio: 0.3
        + Number of iterations of kernels: 2, 2, 2

## Initial estimates

* Cl_pop, Vc_pop, Q_pop, Vp_pop, time0_pop, kt_pop are all set to 1.
* b set to 0.3
* c set to 1 (FIXED)

## Statistical Model & Tasks

### Tasks

#### Estimation of the Population Parameters

* Burn-in-phase (prior to SAEM)
    + Number of iterations: 5
* Exploratory phase
    + Auto-stop criteria: TRUE
    + Maximum number of iterations: 500
    + Minimum number of iterations: 150
    + Stepsize exponent: 0
    + Simulated annealing: TRUE
    + Decreasing rate (variance of the residual errors): 0.95
    + Decreasing rate (variance of the individual parameters): 0.95
* Smooting Phase:
    + Auto stop criteria: TRUE
    + Maximum number of iterations: 200
    + Minimum number of iterations: 50
    + Stepsize exponent: 0.7

#### EBEs

* Conditional Mode - Optimization options
    + Maximum number of iterations: 200
    + Tolerance: .000001
    
#### Conditional distribution

* Conditional distribution (mean and sd) - MCMC convergence assessment
    + Interval length: 50
    + Relative interval length: 0.05
    + Simulated parameters per individual: 10

#### Standard Errors

* Enable task
* Stochastic approximation 
    + Minimum number of iterations: 50
    + Maximum number of iterations: 200
    
#### Likelihood

* Enable task
* Monte Carlo Importance Sampling
    + Monte Carlo size: 10000
    + Degrees of freedom of the t-distribution: Fixed, 5

#### Associated method for scenario

* Use linearization method: FALSE

### Observation model

* Type = Continuous
* Name = DV
* Prediction = centr
* Error Model = proportional
* Distribution = Normal

### Individual model

* Cl, lognormal, Random Effects = TRUE 
* Vc, lognormal, Random Effects = TRUE
* Q, lognormal, Random Effects = TRUE
* Vp, lognormal, Random Effects = TRUE
* time0, lognormal, Random Effects = TRUE
* kt, lognormal, Random Effects = TRUE


No correlations or covariates.

## Run model & save model files

I saved the model files in `monolix-gui-files/`.


# Monolix R API

Next I used all the same settings with the lixoftConnectors package and ran the model again. 

## Setup

* Monolix - 2020R1

```{r}
initializeLixoftConnectors(software = "monolix")
```

* New Project -> Data File -> select "monolix-fake-data.csv"
* ada_model.txt structural model

```{r}
# Monolix needs to know the type of each variable (name)
var_types <- tibble::tribble(
  ~name, ~type,
  "ID", "id",
  "TIME", "time",
  "AMT", "amount",
  "DV", "observation"
)

newProject(
  data = list(
    dataFile = here("data", "monolix-fake-data.csv"),
    headerTypes = var_types$type,
    observationTypes = list(DV = "continuous")
  ),
  modelFile = here("ada_model.txt")
)
```

* Settings -> Project settings
    + Seed: 123456

```{r}
getProjectSettings()$seed
```    

* Settings -> Project settings    
    + Number of chains
        + Automatic choice: TRUE
        + Computed number of chains: 9 
        + Minimum number of individuals: 50

```{r}
setGeneralSettings(nbchains = 9)

getGeneralSettings()
```

* Settings -> Project settings            
    + Transition kernels
        + Target acceptance ratio: 0.3
        + Number of iterations of kernels: 2, 2, 2

```{r}
getMCMCSettings()
```

## Initial estimates

* Cl_pop, Vc_pop, Q_pop, Vp_pop, time0_pop, kt_pop are all set to 1.
* b set to 0.3
* c set to 1 (FIXED)

```{r}
getPopulationParameterInformation()
```

## Statistical Model & Tasks

### Tasks

#### Estimation of the Population Parameters

* Burn-in-phase (prior to SAEM)
    + Number of iterations: 5

```{r}
getPopulationParameterEstimationSettings()$nbburningiterations
```    
    
* Exploratory phase
    + Auto-stop criteria: TRUE
    + Maximum number of iterations: 500
    + Minimum number of iterations: 150
    + Stepsize exponent: 0
    + Simulated annealing: TRUE
    + Decreasing rate (variance of the residual errors): 0.95
    + Decreasing rate (variance of the individual parameters): 0.95

```{r}
getPopulationParameterEstimationSettings()[c(
  "exploratoryautostop", "nbexploratoryiterations", "exploratoryinterval",
  "exploratoryalpha", "simulatedannealing", "tauerrormodel", "tauomega"
  )] 
```
    
* Smooting Phase:
    + Auto stop criteria: TRUE
    + Maximum number of iterations: 200
    + Minimum number of iterations: 50
    + Stepsize exponent: 0.7

```{r}
getPopulationParameterEstimationSettings()[c(
  "smoothingautostop", "nbsmoothingiterations", "smoothinginterval",
  "smoothingalpha"
  )] 
```

#### EBEs

* Conditional Mode - Optimization options
    + Maximum number of iterations: 200
    + Tolerance: .000001

```{r}
getConditionalModeEstimationSettings()
```

    
#### Conditional distribution

* Conditional distribution (mean and sd) - MCMC convergence assessment
    + Interval length: 50
    + Relative interval length: 0.05
    + Simulated parameters per individual: 10

```{r}
getConditionalDistributionSamplingSettings()
```

#### Standard Errors

* Enable task

```{r}
scenario <- getScenario()
scenario$tasks[["standardErrorEstimation"]] <- TRUE
setScenario(scenario)

getScenario()
```

* Stochastic approximation 
    + Minimum number of iterations: 50
    + Maximum number of iterations: 200

```{r}
getStandardErrorEstimationSettings()
```
    
#### Likelihood

* Enable task

```{r}
scenario <- getScenario()
scenario$tasks[["logLikelihoodEstimation"]] <- TRUE
setScenario(scenario)

getScenario()$tasks
```

* Monte Carlo Importance Sampling
    + Monte Carlo size: 10000
    + Degrees of freedom of the t-distribution: Fixed, 5

```{r}
getLogLikelihoodEstimationSettings()
```
#### Associated method for scenario

* Use linearization method: FALSE

```{r}
scenario <- getScenario()
scenario$linearization <- FALSE
setScenario(scenario)

getScenario()$linearization
```

### Observation model

* Type = Continuous
* Name = DV
* Prediction = Cc
* Error Model = proportional
* Distribution = Normal

```{r}
setErrorModel(DV = "proportional")

getContinuousObservationModel()
```

### Individual model

* Cl, lognormal, Random Effects = TRUE 
* Vc, lognormal, Random Effects = TRUE
* Q, lognormal, Random Effects = TRUE
* Vp, lognormal, Random Effects = TRUE
* time0, lognormal, Random Effects = TRUE
* kt, lognormal, Random Effects = TRUE

No correlations or covariates.

```{r}
getIndividualParameterModel()
```

## Run model & save model files

```{r}
runScenario()

saveProject(here("monolix-Rapi-files", "monolix-rapi-example-files.mlxtran"))
```

# Session Info

Adding session info here because the `diffr()` output gets in the way of the session info if it's at the end. 

```{r}
sessionInfo()
```

# Compare Output {.tabset}

Then I compared the files from using the Monolix GUI to the files from using the lixoftConnectors R package. 

## .mlxtran files

```{r}
diffr(
  here("monolix-gui-files", "ada-example-files.mlxtran"),
  here("monolix-Rapi-files", "monolix-rapi-example-files.mlxtran")
)
```

## summary.txt

```{r}
diffr(
  here("monolix-gui-files", "ada-example-files", "summary.txt"),
  here("monolix-Rapi-files", "monolix-rapi-example-files", "summary.txt")
)
```

## populationParameters.txt

```{r}
diffr(
  here("monolix-gui-files", "ada-example-files", "populationParameters.txt"),
  here("monolix-Rapi-files", "monolix-rapi-example-files", "populationParameters.txt")
)
```

## logLikelihood.txt

```{r}
diffr(
  here("monolix-gui-files", "ada-example-files", "LogLikelihood", "logLikelihood.txt"),
  here("monolix-Rapi-files", "monolix-rapi-example-files","LogLikelihood", "logLikelihood.txt")
)
```

## estimatedIndividualParameters.txt

```{r}
diffr(
  here("monolix-gui-files", "ada-example-files", "IndividualParameters", "estimatedIndividualParameters.txt"),
  here("monolix-Rapi-files", "monolix-rapi-example-files", "IndividualParameters", "estimatedIndividualParameters.txt")
)
```

## predictions.txt

```{r}
diffr(
  here("monolix-gui-files", "ada-example-files", "predictions.txt"),
  here("monolix-Rapi-files", "monolix-rapi-example-files", "predictions.txt")
)
```


